---
title: "Weight Lifting Exercise"
author: "Stefan Buchholz"
date: "10. August 2014"
output: html_document
---

#Task-Description
In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: <http://groupware.les.inf.puc-rio.br/har> (see the section on the Weight Lifting Exercise Dataset). 
The aim is to predict the outcome of "classe"-variable with the given variables in the training-dataset.

#Index:
1. Getting Data
2. Data Exploration

#Dataset-Description
The dataset that as downloaded on 08th of August 2014 from the following link: <https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv> describes weight lifting exercises from 6 people. The Dataset contains 19622 observations of 160 variables where the last feature "classes" is an indicator how correct the exercise was done.

The variables in the dataset consists on different datatypes as factors, numerical and integer values. The ranges of the last types are in a negative through a postive range with a lot of NA values.


```{r 1. Download the Datasets}
library(RCurl)

#Windows specific workaround
options(RCurlOptions = list(cainfo = system.file("CurlSSL", "cacert.pem", package = "RCurl")))
#Training data
url <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
website <- getURL(url)
raw_dataset <- read.csv(text = website)

#test data
url_test <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
website <- getURL(url_test)
test <- read.csv(text = website)

#converting columns in testdata
for(i in seq_along(test[,5:159])) {
  test[,i] <- as.numeric(test[,i])
}
```

```{r 2. Data Exploring}
dim(raw_dataset)
colnames(raw_dataset)
summary(raw_dataset)
str(raw_dataset)
head(raw_dataset)
```

```{r exclude mostly NA columns}
# convert columns to get correct NA values
for(i in seq_along(raw_dataset[,5:159])) {
  raw_dataset[,i] <- as.numeric(raw_dataset[,i])
}
y <- NULL
for(i in seq_along(test)) {
  y <- c(y,sum(is.na(test[,i])) > 0)
}

#exclude columns where NAs included
temp <- raw_dataset[,!y]
temp_test <- test[, !y]

#now exclude what doesnt work
data <- temp[, 8:60]
data_test <- temp_test[, 8:60]
```
library(caret)
```{r separating training-set into 2}
intrain <- createDataPartition(y=data$classe, p=0.7, list=FALSE)
training_data <- data[intrain,]
test_data <- data[-intrain,]

```

```{r make Tree and graph it}
library(caret)
set.seed(10)
model <- train(classe ~ ., method="rpart", data=training_data)

library(rattle)
fancyRpartPlot(model$finalModel)
```

```{r predict with tree model}
#predict it
prediction <- predict(model, newdata=test_data)

#calculate missclassification
xtab <- table(prediction, test_data$classe)

confusionMatrix(xtab)
```

```{r model with random forest}
library(caret)
set.seed(10)
trControl = trainControl(method = "cv", number = 4, allowParallel = TRUE,verboseIter = TRUE)
model <- train(classe ~ ., method="rf", traincontrol = trControl, importance=TRUE, data=training_data)

#obove is to slow

model <- randomForest(classe ~ ., importance = TRUE, data = training_data)
varImpPlot(model, scale=TRUE)

#prediction
prediction <- predict(model, newdata=test_data)

#calculate missclassification
xtab <- table(prediction, test_data$classe)
confusionMatrix(xtab)
```

```{r getting most important variables}

```
plot(model, legend=TRUE)
varImpPlot(model, sort=TRUE, n.var=52)$

```{r make predictions for real test data}
prediction <- predict(model, newdata=data_test)
```
answers <- c("B", "A", "B", "A", "A", "E","D","B","A","A","B","C","B","A","E","E","A","B","B","B")

```{r write submission files}

pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}
pml_write_files(answers)

```


